---
title: "INFSCI 2595 Final"
subtitle: "Part Ia: Predictors (categorical)"
author: "Ben Stein"
date: "4/27/2022"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "knit") }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this document

In this document, I perform exploratory data analysis on the categorical predictors (`customer` and `region`).

# Setup

## Libraries and tools

```{r}
library(tidyverse)

source("./tools.R")
```

## Data

```{r}
df <- load_project_data(list(convert_data_types = TRUE, response_to_log = FALSE))
df %>% glimpse()
```

# Exploratory data analysis

## The `customer` predictor

Let's take a look at possible values for the `customer` categorical predictor and their distribution:

```{r, warning=FALSE}
df %>% 
  select(customer) %>%
  ggplot(mapping = aes(x = customer)) + 
  geom_histogram(stat = "count")
```

Key takeaways here are that there aren't many unique values for the predictor (9) and that one category ("Other") contains significantly more observation than others.

## The `region` predictor

Similarly, we can start by looking at the distribution of `region` values.

```{r, warning=FALSE}
df %>% 
  select(region) %>%
  ggplot(mapping = aes(x = region)) + 
  geom_histogram(stat = "count")
```

This factor has even fewer levels (only 3). However, unlike `customer`, the levels are more slightly more similar, with `ZZ` having the most values but the others not terribly far behind.

## Combinations of the categoricals

Let's look at the distribution of these factors together.

**NOTE:** This concept and some of the code to make it happen is borrowed from the airfoil EDA example posted on Canvas for this course.

```{r}
df %>%
  ggplot(mapping = aes(x = customer)) +
  geom_bar(mapping = aes(fill = region), position = "dodge") + 
  scale_fill_viridis_d("region") + 
  theme_bw()
```

This reveals some interesting insights. First, the only level of `customer` that contains every region is `"Other"`, and regions `ZZ` and `YY` never appear in the same level of `customer` except for `"Other"`. Some `customer` levels appear in only one region.

After horsing around with this (and **borrowing more code from the airfoil EDA example**, I realized this can be better explained with a heatmap):

```{r}
df %>% 
  count(customer, region) %>% 
  ggplot(mapping = aes(x = customer, y = region)) +
  geom_tile(mapping = aes(fill = n)) +
  geom_text(mapping = aes(label = n,
                          color = n > median(n)),
            size = 7) +
  scale_fill_viridis_c(guide = 'none') +
  scale_color_manual(guide = 'none',
                     values = c("TRUE" = 'black',
                                "FALSE" = 'white')) +
  theme_bw()
```

This reinforces that customer groups `B`, `D`, `E`, and `K` are less populous and highlights the lack of overlap among customer groups and regions in non-`"Other"` regions. 