---
title: "INFSCI 2595 Final"
subtitle: "Part IIb: Bayesian Regression Models"
author: "Ben Stein"
date: "4/27/2022"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "knit") }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this document

In this document, I evaluate two Bayesian regression models based on findings from the simple frequentist modeling done in the previous section.

# Setup

## Libraries and tools

```{r}
library(tidyverse)
library(rstanarm)

source("./tools.R")
```

## Data

```{r}
df <- load_project_data() %>%
  select(-outcome_numeric)
```


# First Bayesian model (model 3 from the previous section)

Model 3 from the previous section was my best-performing model (RMSE = 0.3299896, R^2 = 0.6233415). It uses all linear predictors. Let's examine this model from a Bayesian perspective.

```{r}
bayesian_model3 <- stan_lm(
  response_log ~ ., 
  data = df,
  prior = R2(location = .2),
  seed = 2022,
  trace = FALSE
)
```

```{r}
bayesian_model3 %>% summary
```

```{r}
plot(bayesian_model3)
```

