---
title: "INFSCI 2595 Final"
subtitle: "Part Ia: Predictors (Lexicon: AFINN)"
author: "Ben Stein"
date: "4/27/2022"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "knit") }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this document

In this document, I perform exploratory data analysis on the variables from the AFINN lexicon data.

# Setup

## Libraries and tools

```{r}
library(tidyverse)

source("./tools.R")
```

## Data

```{r}
df <- load_project_data() %>%
  select(response_log, outcome_numeric, starts_with("xa"))
df %>% glimpse
```

# Distributions

Let's look at the distributions of the variables first.

```{r}
df %>% 
  select(starts_with("xa")) %>%
  summary
```

Note the differing scales of the variables - we'll probably want to center and scale for most models.

```{r}
visualize_distributions(df, "xa")
```

While all variables are loosely gaussian, some are more Gaussian than others (e.g. `xa_02` through `xa_05`).

## Distinct values

As I worked with each group of continuous predictors, I noticed that some seem to have more unique values than others. Let's quantify that:

```{r}
df %>%
  select(starts_with("xa")) %>%
  pivot_longer(cols = everything()) %>%
  group_by(name) %>% 
  unique %>%
  count
```

As we can see here, inputs 1, 4, 7, and 8 have significantly more unique values than the others.

# Correlation within the group

Let's see if any of these inputs are correlated with one another or with the response variable.

```{r}
df %>%
  select(
    starts_with("xa"), 
    response_log
  ) %>%
  cor() %>%
  corrplot::corrplot(type = "upper")
```

We can observe some weak correlations between inputs (e.g. input 1 seems weakly correlated with 2-4 and 7. 4 is correlated with 5, 7, and 8, etc.). Of these, only 1 and 2 seem somewhat correlated with the response variable.

# In relation to the continuous `response` variable

Note that I log-transformed response since that's what I'll be predicting anyway.

```{r}
df %>%
  pivot_longer(cols = starts_with("xa")) %>%
  ggplot(mapping = aes(x = value, y = response_log)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess") + 
  facet_wrap(~ name, scales = "free")
```

It's a bit hard to tell what these distributions really are. Some of them seem as though they might be exponentially distributed, but the loess line looks pretty linear.

# In relation to the `outcome` classification target

```{r}
df %>%
  select(-response_log) %>%
  pivot_longer(cols = starts_with("xa")) %>%
  ggplot(mapping = aes(x = value, y = outcome_numeric)) +
  geom_point(size = 4, alpha = .33) +
  facet_wrap(~ name, scales = "free")
```

It's unclear how this group of predictors relates to `outcome`. High values of the predictors seem to result in more nonevents than events, but it's not clear what happens at middle and low values.
