---
title: "Part iA: Predictors (Lexicon: AFINN)"
subtitle: "INFSCI 2595 Final"
author: "Ben Stein"
date: "4/27/2022"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "knit") }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this document

In this document, I perform exploratory data analysis on the variables from the AFINN lexicon data.

# Setup

## Libraries and tools

```{r}
library(tidyverse)

source("./tools.R")
```

## Data

```{r}
df <- load_project_data() %>%
  select(response_log, outcome_numeric, starts_with("xa"))
df %>% glimpse
```

# Distributions

Let's look at the distributions of the variables first.

```{r}
df %>% 
  select(starts_with("xa")) %>%
  summary
```

Note the differing scales of the variables - we'll probably want to center and scale for most models.

```{r}
visualize_distributions(df, "xa")
```

While all variables are loosely Gaussian, some are more Gaussian than others (e.g. `xa_02` through `xa_05`).

## Distinct values

As I worked with each group of continuous predictors, I noticed that some seem to have more unique values than others. Let's quantify that:

```{r}
df %>%
  select(starts_with("xa")) %>%
  pivot_longer(cols = everything()) %>%
  group_by(name) %>% 
  unique %>%
  count
```

As we can see here, inputs 1, 4, 7, and 8 have significantly more unique values than the others.

# Correlation within the group

Let's see if any of these inputs are correlated with one another.

```{r}
visualize_correlations(df, "xa")
```

There seem to be significant correlations between some of the variables within the group. We'll keep that in mind when selecting variables for modeling.

There is at least moderate correlation among several variables. 

# In relation to the continuous `response` variable

Note that I log-transformed response since that's what I'll be predicting anyway.

```{r}
df %>%
  pivot_longer(cols = starts_with("xa")) %>%
  ggplot(mapping = aes(x = value, y = response_log)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess") + 
  facet_wrap(~ name, scales = "free")
```

It's a bit hard to tell what these distributions really are. Some of them seem as though they might be exponentially distributed, but the loess line looks pretty linear.

# In relation to the `outcome` classification target

```{r}
df %>%
  select(-response_log) %>%
  pivot_longer(cols = starts_with("xa")) %>%
  ggplot(mapping = aes(x = value, y = outcome_numeric)) +
  geom_point(size = 4, alpha = .33) +
  facet_wrap(~ name, scales = "free")
```

It's unclear how this group of predictors relates to `outcome`. High values of the predictors seem to result in more nonevents than events, but it's not clear what happens at middle and low values.
