---
title: "INFSCI 2595 Final"
subtitle: "Part iA: Predictors (Word count features)"
author: "Ben Stein"
date: "4/27/2022"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "knit") }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this document

In this document, I perform exploratory data analysis on the word-count-related inputs.

# Setup

## Libraries and tools

```{r}
library(tidyverse)
source("./tools.R")
```

## Data

```{r}
df <- load_project_data() %>%
  select(response_log, outcome_numeric, starts_with("xw"))

df %>% glimpse
```

# Distributions

Let's look at the distributions of the variables first.

```{r}
df %>% 
  select(-outcome_numeric, -response_log) %>%
  summary
```

Like the sentimentR features, these have more similar scale to one another. 

```{r}
visualize_distributions(df, "xw")
```

In this group, `xw_01` is normally distributed, while the others are more odd. `xw_03` is left-tailed.

## Distinct values

As I worked with each group of continuous predictors, I noticed that some seem to have more unique values than others. Let's quantify that:

```{r}
df %>%
  select(-outcome_numeric, -response_log) %>%
  pivot_longer(cols = everything()) %>%
  group_by(name) %>% 
  unique %>%
  count
```

`xw_01` has more distinct values than the other two by a noticeable amount. Combined with the fact that this variable has the most Gaussian distribution, that suggests that the true distribution may be Gaussian.

# Correlation within the group

Let's see if any of these inputs are correlated with one another or with the response variable.

```{r}
visualize_correlations(df, "xw")
```

`xw_01` is highly correlated with the other two inputs, suggesting we should only use one of these at a time.

# Relationship to target variables

## The continuous `response` variable

Note that I log-transformed response since that's what I'll be predicting anyway.

```{r}
df %>%
  pivot_longer(cols = starts_with("xw")) %>%
  ggplot(mapping = aes(x = value, y = response_log)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess") + 
  facet_wrap(~ name, scales = "free")
```

These relationships look more linear than some of the other predictors, but a little curvy. A low order polynomial might be worth a try.

## The `outcome` classification target

```{r}
df %>%
  select(-response_log) %>%
  pivot_longer(cols = starts_with("xw")) %>%
  ggplot(mapping = aes(x = value, y = outcome_numeric)) +
  geom_point(size = 4, alpha = .33) +
  facet_wrap(~ name, scales = "free")
```

Like the sentimentR group, there's not a clear trend between values of these variables and the `outcome` variable.