---
title: "Part iA: Prediction targets (`outcome`)"
subtitle: "INFSCI 2595 Final"
author: "Ben Stein"
date: "4/27/2022"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "knit") }
  )
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this document

In this document, I perform exploratory data analysis on the `outcome` variable, which is our prediction target for the classification portions of the project.

# Setup

## Libraries and tools

```{r}
library(tidyverse)
source("./tools.R")
```

## Data

```{r}
df <- load_project_data(response_to_log = FALSE, outcome_to_numeric = FALSE)
df %>% glimpse()
```

# Exploratory Data Analysis

The `outcome` is a binary variable which describes whether a given product failed to meet its sales goal (i.e. a value of `"event"` indicates that it **did not** meet its sales goal).

## The variable itself

```{r}
df %>% 
  select(outcome) %>%
  summary()
```


```{r}
df %>% 
  mutate(outcome_logical = ifelse(outcome == "event", 1, 0)) %>%
  select(outcome_logical) %>%
  colMeans
```

The `event` outcome is relatively rare in the dataset (127 times out of `r 127+550` trials, or with about **`r round(127 / (127+550), 3)`** probability). Good for PPG! Or maybe they're not setting their goals ambitiously enough...


## In relation to categorical predictors

Let's see if the distribution is dramatically different when we segment the data using the categorical predictors of `region` and `customer`.

### Region

```{r, warning=FALSE}
df %>%
  select(region, outcome) %>%
  ggplot(mapping = aes(x = outcome)) + 
  geom_histogram(stat = "count") +
  facet_wrap(~region, scales = "free") + 
  theme_bw() +
  theme(axis.text.y = element_blank())
```

It seems as though the event occurs in relatively the same proportion across regions.

### Customer

```{r, warning=FALSE}
df %>%
  select(customer, outcome) %>%
  ggplot(mapping = aes(x = outcome)) + 
  geom_histogram(stat = "count") +
  facet_wrap(~customer, scales = "free") + 
  theme_bw() +
  theme(axis.text.y = element_blank())
```

We start to see some variation here. The `event` outcome seems extremely rare in regions D and E, but closer to the average rate in the rest of the regions. Let's explore a little:

Recall that the overall event probability is about **0.188**. How do the individual region event probabilities compare?

```{r}
df %>%
  select(customer, outcome) %>%
  group_by(customer) %>%
  summarize(
    outcome = mean(ifelse(outcome == "event", 1, 0))
  ) %>%
  ggplot(mapping = aes(x = customer, y = outcome)) +
  geom_col() + 
  geom_hline(yintercept = 0.1875923, color = "red", size = 2, linetype=2)
  
```

This reveals that a few regions (B, D, and E) contribute a much lower event rate to the overall mean event rate.

